### 2024-08-26 visualize_data.Rmd

---
title: "Visualization of Forestry Data"
author: "Tomoki Nishiyama"
date: "2023-10-26"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, arrow, reshape2, ggnetwork, network)
```

# Data
```{r}
forestry_df <- read_parquet("../data/dev/faostat_forestry_trade_long2.parquet")
trade_df <- read_parquet("../data/dev/faostat_forestry_trade_long.parquet")
```

# World Level
```{r}
forestry_df %>% 
  filter(area == "World") %>% 
  ggplot(data=., aes(x=year, y=value, color=item)) +
  geom_line() +
  facet_wrap(facets=~element, ncol = 2, scales = "free")
ggsave(file="../fig/world_item.png", width=20, height=20, dpi=300)
```